Rigidity of Expectations: Additional Evidence from Density Forecasts of Professionals and Households

### 2019-12-17 slides/ExpSlidesnew.tex
slides for brown bag seminar. Will work on following things. 1. true GMM, unconditional moments. 2. randseed lead to SMM.

\documentclass{beamer}
\usepackage{default}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{adjustbox}  % Allows for fitting tables into slide
\usepackage{hyperref}
\usepackage{threeparttable}
\usepackage{caption}
%\usepackage{subcaption}
\usepackage{natbib}

@@

\subsection{Shock-Based Rigidity}

\begin{frame}{Inflation shocks}

\begin{figure}
	\includegraphics[width=0.8\textwidth]{figuresDraft/inf_shocksQ.png} 
\end{figure}

\begin{frame}{Results from \citet{coibion2012can}}

\begin{figure}
	\includegraphics[width=0.9\textwidth]{figuresDraft/RigidityJPEFigure2.png} 
\end{figure}

\begin{frame}{Technology and oil shocks: 1984-2007}

\begin{figure}
	\includegraphics[width=0.5\textwidth,totalheight=0.28\textheight]{figuresDraft/CPIAU_ashocks_nmp_before2007.png}  \\ 
\smallskip
\includegraphics[width=0.5\textwidth,totalheight=0.28\textheight]{figuresDraft/SPFFE_ashocks_nmp_before2007.png} \\
\smallskip
\includegraphics[width=0.5\textwidth,totalheight=0.28\textheight]{figuresDraft/SPFDisg_ab_ashocks_nmp_before2007.png} 
\end{figure}

\begin{frame}{Monetary policy shocks: 1984-2007}

\begin{figure}
	\includegraphics[width=0.5\textwidth,totalheight=0.28\textheight]{figuresDraft/CPIAU_ashocks_before2007.png}  \\
	\smallskip 
	\includegraphics[width=0.5\textwidth,totalheight=0.28\textheight]{figuresDraft/SPFFE_ashocks_before2007.png} \\
	\smallskip
	\includegraphics[width=0.5\textwidth,totalheight=0.28\textheight]{figuresDraft/SPFDisg_ab_ashocks_before2007.png} 
\end{figure}
\end{frame}

\begin{frame}{Technology and oil shocks: 1984-2019}

\begin{figure}
	\includegraphics[width=0.6\textwidth,totalheight=\0.4\textheight]{figuresDraft/Inf_ashocks_nmp.png}  
	\includegraphics[[width=0.6\textwidth,totalheight=\0.4\textheight]{figuresDraft/SPFFE_ashocks_nmp.png}  \\
	\includegraphics[width=0.5\textwidth,totalheight=\0.4\textheight]{figuresDraft/SPFDisg_ab_ashocks_nmp.png} 
	\includegraphics[width=0.5\textwidth,totalheight=\0.4\textheight]{figuresDraft/SPFVar_ab_ashocks_nmp.png}  
\end{figure}
\end{frame}

\begin{frame}{Monetary policy shocks: 1984-2019}

\begin{figure}
	\includegraphics[width=0.5\textwidth,totalheight=\0.4\textheight]{figuresDraft/Inf_ashocks.png} 
	\smallskip
	\includegraphics[width=0.5\textwidth,totalheight=\0.4\textheight]{figuresDraft/SPFFE_ashocks.png} \\
	\smallskip
	\includegraphics[width=0.5\textwidth,totalheight=\0.4\textheight]{figuresDraft/SPFDisg_ab_ashocks.png} 
	\smallskip
	\includegraphics[width=0.5\textwidth,totalheight=\0.4\textheight]{figuresDraft/SPFVar_ab_ashocks.png}  
\end{figure}

@@

\begin{frame}{Sticky Expectation: individual}

For a non-updater since $t-\tau$ ($\tau=0$ for updater),
\begin{itemize}
\item \textbf{Mean} $$E_{i,t}(y_{t+h}|y_{t-\tau}) = \rho^{h+\tau} y_{t-\tau}$$  
\item \textbf{Forecast Error} $$FE_{i,t+h|t} = \underbrace{\sum^{h+\tau}_{s=0} \rho^s \omega_{t+h-s}}_{\substack{\text{weighted sum of} \\ \text{future realized shocks}} }$$
\item \textcolor{blue}{\textbf{Variance}} $$Var_{i,t}(y_{t+h}|y_{t-\tau}) = \sum^{h+\tau}_{s=0}\rho^{2s} \sigma^2_{\omega}$$	
\end{itemize}

\begin{frame}{Sticky Expectation: individual}

$$\text{updater: } \quad \Delta Var_{i,t}(y_{t+h}|y_t)= \sum^{\tau}_{s=0} \rho^{2s}\sigma^2_{\omega}$$
$$\text{non-updater: }  \Delta Var_{i,t|t-\tau-1}(y_{t+h}|y_{t-\tau-1})  = \sigma^2_{\omega}$$

\begin{enumerate}
\item Change in expectation(and variance) depends on if update or not 
\item Cannot observe systematically sluggish response to shocks at individual level 
\end{enumerate}

\begin{frame}{Sticky Expectation: population}
\begin{itemize}
	\item \textbf{Average forecast} 
	\begin{eqnarray*}
		\begin{aligned}
			\bar E_t(y_{t+h}) & = \lambda \underbrace{E_t(y_{t+h})}_{\text{rational expectation at t}} + (1-\lambda) \underbrace{\bar E_{t-1}(y_{t+h})}_{\text{average expectation at } t-1} \\
			& = \lambda E_t(y_{t+h}) + (1-\lambda) (\lambda E_{t-1}(y_{t+h})+ ...) \\
			& =\underbrace{ \lambda \sum^{\infty}_{s=0} (1-\lambda)^s E_{t-s}(y_{t+h})}_{\text{weighted sum of past rational expectations}}
		\end{aligned}
	\end{eqnarray*}
	\item \textbf{Change in average forecast}
	$$\Delta \bar E_t(y_{t+h})=\underbrace{(1-\lambda)}_{\text{stickiness}} \Delta \bar E_{t-1}(y_{t+h}) + \lambda \rho^h \omega_t$$ 
\end{itemize}
\end{frame}

\begin{frame}{Sticky Expectation: population}
\begin{itemize}
\item \textbf{Disagreements}
\begin{eqnarray*}
	\begin{aligned}
		Var_t(y_{t+h} ) & = \lambda \sum^{\infty}_{\tau=0} (1-\lambda)^{\tau} (E_{t|t-\tau}(y_{t+h}) - \bar E_t(y_{t+h}))^2  
	\end{aligned}
\end{eqnarray*}
\item \textbf{Change in disagreements}
$$\Delta Var_t (y_{t+h}) = \rho^{2h} (1-\lambda)\lambda \underbrace{\omega^2_t}_{\text{shock at time t}}$$
\end{itemize}

\begin{enumerate}
\item Disagreements rise after the shock and then gradually decline 
\item Response of disagreements depends on the size of the shock
\end{enumerate}
\end{frame}

\begin{frame}{Sticky Expectation: population}
\begin{itemize}
\item \textcolor{blue}{\textbf{Average variance}}
\begin{eqnarray*}
\begin{aligned}
	\overline {Var}_{t}(y_{t+h}) &  =  (1-\lambda)\underbrace{\overline {Var}_{t-1}(y_{t+h})}_{\text{average variance at t-1}} + \underbrace{\lambda Var_{t}(y_{t+h})}_{\text{variance of updater at t} }
\end{aligned}
\end{eqnarray*}
\item \textbf{Change in average variance}
$$\Delta \overline {Var}_{t}(y_{t+h}) = \underbrace{(1-\lambda)\Delta \overline {Var}_{t-1}(y_{t+h}) - \lambda \rho^{2h}\sigma^2_{\omega}}_{\text{does not depend on shock at t}}$$
\end{itemize}

\begin{enumerate}
\item Average variance does not respond to shocks
\item Average variance has serial correlation with the same rigidity parameter $1-\lambda$ 
\end{enumerate}

\begin{frame}{Noisy Information: individuals}
\begin{itemize}
	\item \textbf{Mean}
	\begin{eqnarray*}
		\begin{aligned}
			E_{i,t}(y_{t+h}) & = \rho^{h}E_{i,t|t}(y_{t}) \\
			E_{i,t|t}(y_{t}) 
			& =  \underbrace{E_{i,t|t-1}(y_{t})}_{\text{prior}} + P \underbrace {(s_{i,t|t}-s_{i,t|t-1})}_{\text{innovations to signals}} \\
			& = (1-PH) E_{i,t|t-1}(y_{t}) + Ps_{i,t} \\
			\text{where }  & P = [P_\epsilon,P_\xi]= \Sigma^y_{i,t|t-1} H(H'\Sigma^y_{i,t|t-1} H + \Sigma^v)^{-1} \\
			\text {where } & \Sigma^y_{i,t|t-1} \text{ is the variance of } y_t \text{ based on prior belief}\\
			\text {and } & \Sigma^v =  \left[ \begin{matrix} 
				\sigma^2_{\epsilon} &  0 \\ 0 & \sigma^2_\xi \end{matrix}\right] 
		\end{aligned}
	\end{eqnarray*}
\end{itemize}
\end{frame}

\begin{frame}{Noisy Information: individuals}
\begin{itemize}
\item \textbf{Change in mean}
\begin{eqnarray*}
	\begin{aligned}
		\Delta E_{i,t|t}(y_{t+h}) & = \underbrace{\rho^h (1-PH)\Delta E_{i,t-1|t-1}(y_{t})}_{\text{Lagged response}} \\
		& + \underbrace{\rho^hPH \Delta y_{i,t} + \rho^h P\Delta v_{i,t}}_{\text{Shocks to signals}}\\
	\end{aligned}
\end{eqnarray*}
\end{itemize}

\begin{enumerate}
	\item Rigidity parameter $1-PH$
	\item Serial correlation at individual level 
	\item Always respond to shocks 
\end{enumerate}

\begin{frame}{Noisy Information: individuals}
\begin{itemize}
	\item \textbf{Variance}

	\begin{eqnarray*}
		\begin{aligned}
			\Sigma^y_{i,t|t} = \Sigma^y_{i,t|t-1} - \Sigma^y_{i,t|t-1} H'(H \Sigma^y_{i,t-1} H' +\Sigma^v) ^{-1}H \Sigma^y_{i,t|t-1} 
		\end{aligned}
	\end{eqnarray*}
	\item \textbf{Change in variance} 
	$$\Delta \Sigma^y_{i,t|t} <0$$

\end{itemize}
\begin{enumerate}
	\item It does not depend on the realizations of the signal. 
	\item It decreases unambiguously from $t-1$ to $t$. 
	\item The two properties carry through to h-period ahead forecast 
\end{enumerate}

\begin{frame}{Noisy Information: population}
\begin{itemize}
\item \textbf{Mean}

\begin{eqnarray*}
	\begin{aligned}
		\bar E_{t|t} (y_{t+h}) & = \rho^h [(1-PH) \underbrace{\bar E_{t-1}(y_{t+h})}_{\text{Average prior}} + P \underbrace{\bar s_{t}}_{\text{Average Signals}}] \\
		& = (1-PH) \bar E_{t-1}(y_{t+h}) + P [\epsilon_t, 0]' \\
		& = (1-PH) \bar E_{t-1}(y_{t+h}) + P \epsilon_t
	\end{aligned}
\end{eqnarray*}
\end{itemize}

\begin{enumerate}
	\item Same properties to the individual forecast 
\end{enumerate}

\begin{frame}{Noisy Information: population}
\begin{itemize}
\item \textbf{Disagreements}
\begin{eqnarray*}
\begin{aligned}
	Var_t(y_{t+h}) & = E((E_{i,t|t}(y_{t+h}) - \bar E_t(y_{t+h}))^2) \\
	& = \rho^{2h} P^2_\xi \sigma^2_\xi  
\end{aligned}
\end{eqnarray*}
\end{itemize}

\begin{enumerate}
	\item increase with the forecast horizon
	\item depends on noisiness private signals, but not on that of public signals and the variance of the true variable $y$ 
	\item increase with the rigidity parameter $P$ in this model
\end{enumerate}

\begin{frame}{Noisy Information: population}
\begin{itemize}
\item \textbf{Change in disagreements}
\begin{eqnarray*}
\begin{aligned}
\Delta Var_t(y_{t+h}) & = \rho^{2h}(1-\rho^2) P^2_\xi \sigma^2_\xi >0
\end{aligned}
\end{eqnarray*}
\end{itemize}

\begin{enumerate}
	\item disagreements increase as time goes from $t-1$ to $t$. 
	\item disagreements increase as approaching the variable of forecast
\end{enumerate}

\begin{frame}{Noisy Information: population}
\begin{itemize}
\item \textbf{Average variance}
\begin{eqnarray*}
\begin{aligned}
\bar Var_t (y_{t+h}) = \bar \Sigma^y_t
\end{aligned}
\end{eqnarray*}
\item \textbf{Change in average variance}
\begin{eqnarray*}
\Delta Var_t(y_{t+h}) < 0 
\end{eqnarray*}

\end{itemize}

\begin{enumerate}
	\item average variance is the same as individual variance, not depend on signals
	\item the variance unambiguously drop over time 
\end{enumerate}

>>>>

\documentclass{beamer}
\usepackage{default}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{adjustbox}  % Allows for fitting tables into slide
\usepackage{hyperref}
\usepackage{threeparttable}
\usepackage{caption}
%\usepackage{subcaption}
\usepackage{natbib}
\usepackage{adjustbox}

@@

\begin{frame}{Estimation: minimum distance approach}
	\begin{itemize}
\item model-specific moment dynamics as a function real time data and process parameters 
\item minimize the distance (percent deviation) from the data moments to find the parameter of expectation formation	
	\end{itemize}
\end{frame}

\begin{frame}{Results: professionals and SEAR}
\begin{figure}[ht]
	\centering
    \label{SE_diag_SPF}
		\includegraphics[width=0.19\textwidth, height = \0.95\textheight]{figures/spf_se_est_diag0.png}
		\includegraphics[width=0.19\textwidth, height = \0.95\textheight]{figures/spf_se_est_diag1.png}
		\includegraphics[width=0.19\textwidth, height = \0.95\textheight]{figures/spf_se_est_diag2.png}
		\includegraphics[width=0.19\textwidth, height = \0.95\textheight]{figures/spf_se_est_diag3.png}
		\includegraphics[width=0.19\textwidth, height = \0.95\textheight]{figures/spf_se_est_diag4.png}
\end{figure}
\end{frame}

\begin{frame}{Results: households and SEAR}

	\begin{frame}{Results: households and SESV}
		\begin{figure}[ht]
			\centering
			\label{SESV_diag_SCE}
			\includegraphics[width=0.19\textwidth, height = \0.95\textheight]{figures/sce_se_est_sv_diag0.png}
			\includegraphics[width=0.19\textwidth, height = \0.95\textheight]{figures/sce_se_est_sv_diag1.png}
			\includegraphics[width=0.19\textwidth, height = \0.95\textheight]{figures/sce_se_est_sv_diag2.png}
			\includegraphics[width=0.19\textwidth, height = \0.95\textheight]{figures/sce_se_est_sv_diag3.png}
			\includegraphics[width=0.19\textwidth, height = \0.95\textheight]{figures/spf_se_est_sv_diag4.png}
		\end{figure}
	\end{frame}

	\begin{frame}{SEAR parameters}
		\begin{figure}[ht]
			\centering
			\label{SE_diag_SCE}
			\includegraphics[width=0.19\textwidth, height = \0.95\textheight]{figures/sce_se_est_diag0.png}
			\includegraphics[width=0.19\textwidth, height = \0.95\textheight]{figures/sce_se_est_diag1.png}
			\includegraphics[width=0.19\textwidth, height = \0.95\textheight]{figures/sce_se_est_diag2.png}
			\includegraphics[width=0.19\textwidth, height = \0.95\textheight]{figures/sce_se_est_diag3.png}
			\includegraphics[width=0.19\textwidth, height = \0.95\textheight]{figures/sce_se_est_diag4.png}
		\end{figure}
	\end{frame}

	\begin{frame}{SESV parameters}
		\begin{table}
			\centering
			\caption{GMM Estimates of Parameters of SESV}
			\label{GMM_Est_SE_SV_Table}
			\begin{tabular}{llllll}
				\hline 
				0        & 1  & 2    & 3   & SE: $\hat\lambda_{SPF}$(Q) & SE: $\hat\lambda_{SCE}$(M) \\
				\hline 
				Forecast &    &      &     & 0.23                       & 0.01                       \\
				FE       &    &      &     & 0.12                       & 1                          \\
				Forecast & FE &      &     & 0.12                       & 1                          \\
				Forecast & FE & Disg &     & 0.12                       & 1                          \\
				Forecast & FE & Disg & Var & 0.12                       & 1           \\
				\hline               
			\end{tabular}
		\end{table}	
	\end{frame}

	\begin{frame}{SESV parameters}
		\begin{table}
			\centering
			\caption{GMM Estimates of Parameters of SE and Inflation Process}
			\label{GMM_Est_SE_Table}
			\begin{tabular}{llllllllllll}
				\hline 
				& 1  & 2    & 3   & SE: $\hat\lambda_{SPF}$(Q) & SE: $\hat\lambda_{SPF}$(Q) & SE: $\rho$ & SE: $\sigma$ & SE: $\hat\lambda_{SCE}$(M) & SE: $\hat\lambda_{SCE}$(M) & SE: $\rho$ & SE: $\sigma$ \\
				\hline 
				Forecast &    &      &     & 0.05                       & 0.03                       & 1          & 0.1          & 1                          & 1                          & 1          & 0.1          \\
				FE       &    &      &     & 0.05                       & 0.03                       & 1          & 0.1          & 1                          & 1                          & 1          & 0.1          \\
				Forecast & FE &      &     & 0.05                       & 0.03                       & 1          & 0.1          & 1                          & 1                          & 1          & 0.1          \\
				Forecast & FE & Disg &     & 0.7                        & 0.38                       & 0.88       & 0.1          & 0.02                       & 0.03                       & 0.98       & 0.1          \\
				Forecast & FE & Disg & Var & 0.7                        & 0.38                       & 0.88       & 0.2          & 0.01                       & 0.03                       & 0.98       & 1.11     \\
				\hline    
			\end{tabular}
		\end{table}
\end{frame}

\begin{frame}{Conclusion}
\begin{itemize}
	\item Sticky expectation(SE) matches data of inflation and expectations better compared to nosiy information (NI) 
	\item Within model,  households are more irrational compared professionals
	\item Incorporating higher moments, i.e. uncertainty helps ``discipline'' theories on expectation formation
	\item  Higher moments from surveys also contain useful information about the inflation dynamics itself
\end{itemize}	
\end{frame}

@@
